---
title: "Bellabeat Case Study Analysis Using R"
author: "Derrick"
date: "2023-09-29"
output:
  html_document:
    df_print: paged
---

## Bellabeat CaseStudy

### Bellabeat is a high-tech manufacturer of health-focused products for women.

### Business Task:

##### To focus on a Bellabeat product and analyze smart device usage data in order to gain insight into how people are already using their smart devices. Then, using this information, make high-level recommendations for how these trends can inform Bellabeat marketing strategy.

##### To analyze smart device usage data in order to gain insight into how consumers use non-Bellabeat smart devices. Then, select one Bellabeat product to apply these insights to in your presentation

### Data Sources:

##### Kaggle: Kaggle is the world's largest data science community with powerful tools and resources to help one achieve their data science goals. The Dataset used is the FitBit Fitness Tracker Data.

### Stakeholders:

##### (i.) Urška Sršen

##### (ii.) Sando Mur

##### (iii.) Bellabeat marketing analytics team

```{r Setting up: Loading packages}
install.packages("tidyverse")
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Getting out the datasets

```{r importing the dataset}
daily_Act <- read_csv("dailyActivity_merged.csv")
hourly_Cal <- read_csv("hourlyCalories_merged.csv")
Intensities_22 <- read_csv("dailyIntensities_merged.csv")
sleep <- read_csv("sleepDay_merged.csv")
weight <- read_csv("weightLogInfo_merged.csv")
```

## Aligning the formatting

```{r}
# Aligning the formatting
hourly_Cal$ActivityHour=as.POSIXct(hourly_Cal$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
hourly_Cal$time <- format(hourly_Cal$ActivityHour, format = "%H:%M:%S")
hourly_Cal$date <- format(hourly_Cal$ActivityHour, format = "%m/%d/%y")
# activity
daily_Act$ActivityDate=as.POSIXct(daily_Act$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
daily_Act$date <- format(daily_Act$ActivityDate, format = "%m/%d/%y")
# sleep
sleep$SleepDay=as.POSIXct(sleep$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep$date <- format(sleep$SleepDay, format = "%m/%d/%y")
```

## Seeing the Unique Ids

```{r, Seeing the unique Ids}
n_distinct(hourly_Cal$Id)
n_distinct(daily_Act$Id)
n_distinct(Intensities_22$Id)
n_distinct(sleep$Id)
n_distinct(weight$Id)
```

#Summarizing the data

```{r, activity}
daily_Act %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, Calories) %>%
  summary()

# explore num of active minutes per category
daily_Act %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>%
  summary()

# calories
hourly_Cal %>%
  select(Calories) %>%
  summary()

# sleep
sleep %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>%
  summary()

# weight
weight %>%
  select(WeightKg, BMI) %>%
  summary()
```

#Merging the data

```{r, merged Data}
merged_Data <- merge(sleep, daily_Act, by=c('Id', 'date'))
head(merged_Data)
```

# Visualization of the Data

```{r, visuals}
ggplot(data=daily_Act, aes(x=TotalSteps, y=Calories)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Calories")
```

#### The above scatterplot clearly shows the relationship between the Total Steps and the calories burned. 
#### It is evident that the higher the *total steps* the higher the amount of calories burned up. 

# Further visuals

```{r, more visuals}
ggplot(data=sleep, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + 
  geom_point(color=red)+ labs(title="Total Minutes Asleep vs. Total Time in Bed")
```
#### There is a linear relationship between Total Minutes Asleep and Total Time in Bed. This means Apps notifications should be employed to notify its users on when to sleep


# Relationship between Total Minutes Asleep and Sedentry Minutes

```{r, Relationship}
ggplot(data=merged_Data, aes(x=TotalMinutesAsleep, y=SedentaryMinutes)) + 
  geom_point(color='darkblue') + geom_smooth() +
  labs(title="Minutes Asleep vs. Sedentary Minutes")
```

#### The relationship between the sedentary minutes and the total minutes asleep is negative. But this correlation does not necessary mean causation.
#### A recommendation would be for the Bellabeat App to use a notification message to its users to help reduce the sedentary minutes.

# Recommendation to the Stakeholders

#### 1. Bellabeat App to encourage users who want to lose weight by monitoring their daily calorie consumption. Bellabeat to suggest some ideas for low-calorie lunch and dinner.

#### 2. To improve users'sleep time, Bellabeat should consider using app notifications to go to bed.

#### 3. Most active times are between 5 pm and 7 pm as per the analysis. Bellabeat to have systems of reminding their users on being active by having their daily run or walk.

#### 4.The Bellabeat app to notify their users on reducing sedentary time in order to have improved sleep times

